---
title: "Data ingestion for 2017"
author: "K Todd-Brown (ktoddbrown@gmail.com)"
date: "11/13/2017"
output:
  pdf_document:
    fig_caption: yes
    toc: yes
  html_document: default
---

```{r setup}
library(SoilDataR) #library(devtools); install_github("ktoddbrown/soilDataR")
library(ggplot2) #make pretty plots
library(dplyr)
library(knitr)
library(lubridate)
library(tidyr)

#mapping librarys to help with global/regional plots
library(ggmap)
library(maps)
library(mapdata)
library(fiftystater)

```

Datasets are:
1) Treat: Peat properties synthesis dataset (2MB, XLSX format, download only; ISCNtemplate_Treat_peatProps_v2): This dataset is a synthesis of literature and site-level data on peat properties, C, N, 14C, and vegetation from 366 sites worldwide. Data are available for nearly 16,000 layers from 659 profiles. Data contributed by Claire Treat.
2) Alamos soil C stocks (<1MB, XLSX format, download only; ISCNtemplate_Alamos): This site-level dataset comes courtesy of Kris Johnson and collaborators at ITSON (Obregon, MX). It contains 30 profiles sampled by quantitative pit as part of a NASA-supported C monitoring study.
3) Berhe et al 2012. Fractionation example from the Power Center Working Group, manuscript DOI: 10.1029/2011JG001790

```{r loadData}
##source('../../SoilDataR/R/processData_Templet.R') ##Uncomment to debug template files
ingestFiles <- list(
  filename = c('../repoData/Treat_2015/ISCNtemplate_Treat_peatProps_v2.xlsx',
               '../repoData/Alamos/ISCNtemplate_Alamos.xlsx',
               '../repoData/Berhe2012/Berhe_2012.xlsx'),
  keyFile = c(rep('../templates/ISCNtemplate_2016Key.xlsx', 2),
              '../templates/PowellCenterKey.xlsx'), 
  verticalSheets = c(rep('metadata', 2), ''),
  skip=list(c(1:2), 1:2, NA))

data.ls <- list(study=data.frame(), field=data.frame(), sample=data.frame(), treatment=data.frame())
for(ii in 1:length(ingestFiles$filename)){
  temp <- processData_Templet(
    filename=ingestFiles$filename[[ii]],
    key.df=readxl::read_excel(path=ingestFiles$keyFile[[ii]], sheet='headerKey'), 
    skip=ingestFiles$skip[[ii]], 
    verticalSheets=ingestFiles$verticalSheets[[ii]])
  
  ##append data set name
  datasetName <- unique(temp$study$dataset_name[!is.na(temp$study$dataset_name)])
  temp$field$dataset_name <- datasetName
  temp$sample$dataset_name <- datasetName
  if(nrow(temp$treatment) > 0) temp$treatment$dataset_name <- datasetName
  
  ##append units to sample
  temp$sample <- temp$sample %>%
    mutate(unit = as.character(unit)) %>%
    left_join(select(temp$key %>% filter(type == 'value'), var, hardUnit)) %>%
    mutate(var=as.factor(var))
  
  data.ls$study <- bind_rows(data.ls$study,
                             temp$study)
  data.ls$field <- bind_rows(data.ls$field,
                             temp$field)
  data.ls$sample <- bind_rows(data.ls$sample,
                              temp$sample)
  data.ls$treatment <- bind_rows(data.ls$treatment,
                                 temp$treatment)
}

##Filter the messy study names
data.ls$study <- data.ls$study %>% 
  filter(!is.na(dataset_name) & !is.na(curator_email)) %>%
  arrange(dataset_name)

data.ls$sample <- data.ls$sample %>%
  mutate(unit = if_else(is.na(unit), hardUnit, unit))
```

#Observation time
```{r obsTime}
location.df <- data.ls$field %>% 
  select(lat, long, observation_date, state, country, dataset_name) %>%
  unique() %>%
  ###cetch any N|S or E|W notations
  mutate(lat = if_else(grepl('S', as.character(lat)), -1*as.numeric(gsub('S', '', as.character(lat))),
                                       as.numeric(gsub('N', '', as.character(lat)))),
         long = if_else(grepl('W', as.character(long)), -1*as.numeric(gsub('W', '', as.character(long))),
                                as.numeric(gsub('E', '', as.character(long))))) %>%
  ###convert the observation dates
  mutate(observation_date = as.character(observation_date)) %>%
  separate(observation_date, c('monthStr', 'dayStr', 'yearStr'), 
           remove=FALSE, fill='left') %>%
  mutate(year=if_else(is.na(as.numeric(yearStr)), as.numeric(yearStr),
                      if_else(as.numeric(yearStr) < 20, as.numeric(yearStr) + 2000,
                              if_else(as.numeric(yearStr) < 100, as.numeric(yearStr) + 1900,
                                      as.numeric(yearStr)))),
         month=if_else(is.na(as.numeric(monthStr)), 1,  as.numeric(monthStr)),
         day=if_else(is.na(as.numeric(dayStr)), 1, as.numeric(dayStr))) %>%
  select(-contains('Str')) %>% 
  mutate(obsDate = ymd(paste(year, month, day, sep='-'))) %>%
  arrange(lat, long, obsDate) %>%
  ###segment everything
  mutate(yrCut = cut(year, seq(from = floor(min(year, na.rm=TRUE)/10)*10, 
                               to = ceiling(max(year, na.rm=TRUE)/10)*10, by=10),
                     dig.lab=4),
         latCut = cut(lat, seq(-90, 90, by=0.05)),
         longCut = cut(long, seq(-180, 180, by=0.05))) %>%
  ###replace common country names
  mutate(country = if_else(grepl('USA', country), 'United States', country))

timeSpaceCounts <- location.df %>%
  group_by(yrCut,  country) %>%
  tally

print(timeSpaceCounts)
```
```{r plotTime}
ggplot(location.df, aes(x=obsDate, fill=dataset_name)) + geom_histogram()

mapWorld <- borders("world", colour="gray80", fill="gray80") # create a layer of borders
#ggplot() + mapWorld
ggplot(location.df) +
  mapWorld + 
  #geom_hex(aes(x=long, y=lat), bins=200) + 
  geom_point(aes(x=long, y=lat, color=dataset_name)) +
  scale_fill_gradient(trans='log10') +
  theme_bw() +
  theme(text=element_text(size=18),
        legend.text=element_text(size=10),
        axis.title=element_blank()) +
  #ylim(45, 90) + 
  #coord_map(projection='azequidistant') +
  facet_wrap(~yrCut)

```

# Site locations
## Lat-lon map
```{r mapSites}
mapWorld <- borders("world", colour="gray80", fill="gray80") # create a layer of borders
#ggplot() + mapWorld
ggplot(unique(location.df[, c('lat', 'long')])) +
  mapWorld + 
  geom_hex(aes(x=long, y=lat), bins=200) + 
  scale_fill_gradient(trans='log10') +
  theme_bw() +
  theme(text=element_text(size=18)) +
  labs(x='', y='')
```

# Measruements distribution
```{r measurementHist, fig.height=10, fig.width=7}
ggplot(data.ls$sample) +
  geom_histogram(aes(x=value, fill=dataset_name)) +
  facet_wrap(var~unit, scales='free', ncol=2)
```

# Comparison with ISCN3

```{r loadISCN3}
#source('../../SoilDataR/R/processData_ISCN3.R')
if(file.exists('../temp/ISCN3.RData')){
  load('../temp/ISCN3.RData')
  
  
}else{
  ISCN3 <- processData_ISCN3(layersDir='../repoData/ISCN_3/Layers/', metaDir='../repoData/ISCN_3/Meta/',
                             keyFile='../repoData/ISCN_3/ISCNKey.xlsx',
                             loadVars=as.character(unique(data.ls$sample$var)))
}
```

```{r ISCNclean}
##Add the hard coded units
hardUnits.df <- ISCN3$ISCNKey %>% filter(type == 'value', !is.na(hardUnit)) %>% select(var, hardUnit)
varOfInterest <- as.character(unique(data.ls$sample$var))
ISCN3$measure <- ISCN3$measure %>% 
  filter(var %in% varOfInterest) %>% #only pull variables we care about
  left_join(select(filter(ISCN3$ISCNKey, type=='value'), var, hardUnit), by='var') %>% #add in hard units
  mutate(unit=if_else(grepl('\\w', unit) , as.character(unit), hardUnit))

measureID_valid <- ISCN3$measure$measureID ##too many df with the same variables
ISCN3$sample <- ISCN3$sample %>% filter(measureID %in% measureID_valid) #only keep samples with variables of interest

fieldID_valid <- unique(ISCN3$sample$fieldID)
ISCN3$field <- ISCN3$field %>% 
  filter(fieldID %in% fieldID_valid) %>% #only keep fields with variables of interests
  mutate(lat=as.numeric(lat), lon=as.numeric(lon)) #convert lat-lon to numbers

```

```{r}
ISCNLocation <- ISCN3$field %>% select(ends_with('_name'), lat, lon, state, country, observation_date) %>%
  rename(long=lon) %>% 
  select(lat, long, observation_date, state, country, dataset_name) %>%
  unique() %>%
  ##Convert to numeric
  mutate(dataset_name='ISCN3', long=as.numeric(long), lat=as.numeric(lat)) %>%
  mutate(observation_date = as.character(observation_date)) %>%
  separate(observation_date, c('monthStr', 'dayStr', 'yearStr'), 
           remove=FALSE, fill='left') %>%
  mutate(year=if_else(is.na(as.numeric(yearStr)), as.numeric(yearStr),
                      if_else(as.numeric(yearStr) < 20, as.numeric(yearStr) + 2000,
                              if_else(as.numeric(yearStr) < 100, as.numeric(yearStr) + 1900,
                                      as.numeric(yearStr)))),
         month=if_else(is.na(as.numeric(monthStr)), 1,  as.numeric(monthStr)),
         day=if_else(is.na(as.numeric(dayStr)), 1, as.numeric(dayStr))) %>%
  select(-contains('Str')) %>% 
  mutate(obsDate = ymd(paste(year, month, day, sep='-')))
         
allDatasets <- c(unique(as.character(ISCNLocation$dataset_name)), unique(as.character(location.df$dataset_name)))
assert_that(length(allDatasets) == length(unique(allDatasets)))
            

allLocations <- location.df %>%
  bind_rows(ISCNLocation) %>%
  mutate(yrCut = cut(year, seq(from = floor(min(year, na.rm=TRUE)/10)*10, 
                               to = ceiling(max(year, na.rm=TRUE)/10)*10, by=10),
                     dig.lab=4),
         latCut = cut(lat, seq(-90, 90, by=0.05)),
         longCut = cut(long, seq(-180, 180, by=0.05)))

allLocations %>% group_by(dataset_name) %>% 
  summarize(minYear = min(year, na.rm=TRUE), maxYear=max(year, na.rm=TRUE), x=length(year))
```

```{r}
mapWorld <- borders("world", colour="gray80", fill="gray80") # create a layer of borders
#ggplot() + mapWorld
ggplot(allLocations) +
  mapWorld + 
  #geom_hex(aes(x=long, y=lat), bins=200) + 
  geom_hex(data=filter(allLocations, dataset_name == 'ISCN3'), aes(x=long, y=lat), bins=200) +
  geom_point(data=filter(allLocations, dataset_name != 'ISCN3'), aes(x=long, y=lat, color=dataset_name)) +
  scale_fill_gradient(trans='log10', low='gold', high='gold4') +
  theme_bw() +
  theme(text=element_text(size=18),
        legend.text=element_text(size=10),
        axis.title=element_blank(),
        legend.position='bottom') +
  #ylim(45, 90) + 
  #coord_map(projection='azequidistant') +
  facet_wrap(~yrCut)
```

```{r}
refData <- ISCN3$measure %>%
  left_join(ISCN3$sample)

ggplot(data.ls$sample %>% filter(var %in% c('14c_age', 'n_tot', 'c_tot', 'loi'))) +
  geom_histogram(data=refData %>% filter(var %in% c('14c_age', 'n_tot', 'c_tot', 'loi')),
                 aes(x=value, y=..density..), fill='grey') + 
  geom_histogram(aes(x=value, y=..density.., fill=dataset_name), alpha=0.5) +
  facet_wrap(var~unit, scales='free')

ggplot(data.ls$sample %>% filter(!var %in% c('14c_age', 'n_tot', 'c_tot', 'loi'))) +
  geom_histogram(data=refData %>% filter(!var %in% c('14c_age', 'n_tot', 'c_tot', 'loi')), 
                 aes(x=value, y=..density..), fill='grey') + 
  geom_histogram(aes(x=value, y=..density.., fill=dataset_name), alpha=0.5) +
  scale_x_log10() +
  facet_wrap(var~unit, scales='free')

```
