---
title: "ISCN3 visualization and report"
author: "K Todd-Brown (ktoddbrown@gmail.com)"
date: "5/26/2017"
output:
  pdf_document:
    fig_caption: yes
    toc: yes
  html_document: default
---

The International Soil Carbon Network version 3 data base (ISCN3) is available [here](http://iscn.fluxdata.org/).
Currently the Layers data is ingestion and summarized below (work is underway to also bring in profile level data). 
Much of the data is from [NRCS](https://www.nrcs.usda.gov/wps/portal/nrcs/site/soils/home/) (Sept 2014 version of the soil characterization data base) and methods can be refered to in SSL (SoiL Survey Manual, R Burt) manual. 

```{r setup}
library(SoilDataR) #library(devtools); install_github("ktoddbrown/soilDataR")
library(ggplot2) #make pretty plots
library(plyr) #data management
library(dplyr)
library(knitr)
library(lubridate)

#mapping librarys to help with global/regional plots
library(ggmap)
library(maps)
library(mapdata)
library(fiftystater)
```

```{r loadData, cache=TRUE}
ISCN3 <- SoilDataR::processData_ISCN3(dir='../repoData/ISCN_3/')
```


#Observation time
```{r obsTime}
location.df <- unique(ISCN3$field[, c('lat', 'long', 'observation_date', 
                                       'state', 'country')])
location.df$observation_date <- as.character(location.df$observation_date)

location.df <- ddply(location.df, c('observation_date'), function(xx){
  temp <- unlist(strsplit(xx$observation_date[1], '/'))
  xx$month <- NA
  xx$day <- NA
  xx$year <- NA
  if(length(temp) > 1){
    xx$month <- as.numeric(temp[1])
    xx$day <- as.numeric(temp[2])
    xx$year <- ifelse(as.numeric(temp[3]) > 100, as.numeric(temp[3]), 
                      ifelse(as.numeric(temp[3]) < 20, as.numeric(temp[3]) + 2000,
                             as.numeric(temp[3]) + 1900))
    
  }else{
    xx$year <- as.numeric(temp[1])
    if(is.finite(xx$year[1])){
      xx$month <- 1
      xx$day <- 1
    }
  }
  
  return(xx)
})

location.df$obsDate <- ymd(paste(location.df$year, location.df$month, location.df$day, sep="-"))
location.df <- location.df[order(location.df$lat, location.df$long, location.df$obsDate),]
location.df$yrCut <- cut(location.df$year, c(1928, 1940, 1950, 1960, 1970, 1980, 1990, 2000, 2014), dig.lab=4)
location.df$latCut <- cut(location.df$lat, seq(-90, 90, by=1))
location.df$longCut <- cut(location.df$long, seq(-180, 180, by=1))

timeSpaceCounts <- location.df %>%
  group_by(yrCut, latCut, longCut) %>%
  tally
```
```{r plotTime}
ggplot(location.df, aes(x=obsDate)) + geom_histogram()

mapWorld <- borders("world", colour="gray80", fill="gray80") # create a layer of borders
#ggplot() + mapWorld
ggplot(location.df) +
  mapWorld + 
  geom_hex(aes(x=long, y=lat), bins=200) + 
  scale_fill_gradient(trans='log10') +
  theme_bw() +
  theme(text=element_text(size=18)) +
  labs(x='', y='') +
  facet_wrap(~yrCut)

ggplot(location.df, aes(x=long, y=lat)) + geom_point(alpha=0.1)

ggplot(timeSpaceCounts, aes(x=n)) + geom_histogram() + scale_x_log10()
```

# Site locations
## Lat-lon map
```{r mapSites}

mapWorld <- borders("world", colour="gray80", fill="gray80") # create a layer of borders
#ggplot() + mapWorld
ggplot(unique(ISCN3$field[, c('lat', 'long')])) +
  mapWorld + 
  geom_hex(aes(x=long, y=lat), bins=200) + 
  scale_fill_gradient(trans='log10') +
  theme_bw() +
  theme(text=element_text(size=18)) +
  labs(x='', y='')
```

## Sites with states but no lat-lon
```{r statemap}
noLatLon <- unique(subset(ISCN3$field, is.na(lat+long))$fieldID)

stateCounts <- ddply(subset(ISCN3$field, is.na(lat+long)), c('state'), 
                     summarize, count=length(unique(fieldID)))
stateCounts$state <- tolower(stateCounts$state)

cat('Following are not considered states:\n')
kable(subset(stateCounts, ! state %in% fifty_states$id))
stateCounts <- subset(stateCounts, state %in% fifty_states$id)

ggplot(stateCounts, aes(map_id = state)) +
  geom_map(aes(fill=count), map=fifty_states) +
  expand_limits(x = fifty_states$long, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  scale_fill_gradient(trans='log10') +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", 
        panel.background = element_blank())
```

## Sites with country but no lat-lon
```{r countryMap}
countryCounts <- ddply(subset(ISCN3$field, is.na(lat+long)), 'country',  summarize, count=length(unique(fieldID)))
names(countryCounts)[1] <- 'region'
countryCounts$region <- as.character(countryCounts$region)
countryCounts$region[grepl('United States', countryCounts$region)] <- 'USA'

map.world <- map_data(map="world")

cat('These countries are unknown to the map\n')
kable(subset(countryCounts, ! region %in% map.world$region))
missingCountries <- data.frame(region=setdiff(map.world$region, countryCounts$value), count=NA)
countryCounts <- rbind(countryCounts, missingCountries)
#map.world <- merge(map.world, countryCounts, all=TRUE)
ggplot(countryCounts, aes(map_id = region)) +
  geom_map(aes(fill=count), map=map.world) +
  expand_limits(x = map.world$long, y = map.world$lat) +
  coord_cartesian(xlim=c(-180,180), ylim=c(-90,90)) +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  scale_fill_gradient(trans='log10') +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", 
        panel.background = element_blank())
```

`r nrow(unique(subset(ISCN3$field, (country=='Unknown' | country == '') & is.na(lat+long), select=c('fieldID'))))` sites with no known country nor lat-lon.

## Site count by country map (include sites w/ lat-lon)
```{r totCountryCount}
countryCounts <- ddply(ISCN3$field, 'country', summarize, count=length(unique(fieldID)))
names(countryCounts)[1] <- 'region'
countryCounts$region <- as.character(countryCounts$region)
countryCounts$region[grepl('United States', countryCounts$region)] <- 'USA'

map.world <- map_data(map="world")

missingCountries <- data.frame(region=setdiff(map.world$region, countryCounts$region), count=NA)
setdiff(countryCounts$region, map.world$region)
countryCounts <- rbind(countryCounts, missingCountries)
#map.world <- merge(map.world, countryCounts, all=TRUE)
ggplot(countryCounts, aes(map_id = region)) +
  geom_map(aes(fill=count), map=map.world) +
  expand_limits(x = map.world$long, y = map.world$lat) +
  coord_cartesian(xlim=c(-180,180), ylim=c(-90,90)) +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  scale_fill_gradient(trans='log10') +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", 
        panel.background = element_blank())
```

# Summary of measurements

The following variables are listed in 
```{r types}
table(ISCN3$measurement$type)
```
All measurements are on oven dry soils and in mass percent where percent noted.

## Soil organic carbon stock

+ *soc* calculated soil carbon stock of layer/profile

```{r SOC}
ggplot(merge(subset(ISCN3$sample, grepl('^soc', measurementID) & value > 0),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit)

summary(subset(ISCN3$sample, grepl('^soc', measurementID) & (value < 0)))
```

## Bulk density

+ *bd_samp* is the bulk density of the <2mm (fine earth) fraction, in which the mass is expressed on an oven-dry (105 deg) basis. This is the best form of Db for the most uses.
+ *bd_tot* is the whole soil bulk density, includes fine earth fraction and rocks.
+ *bd_whole* is the bulk density of the whole soil (coarse frags, if present, included), expressed on a 1/3 bar moisture content (a little drier than typical field moist, but not oven-dry). It is collected by the clod method. For a variety of reasons, including the presence of water, possible coarse frags, and the tendency of clods to cleave along pore spaces, this will overestimate Db relative to bd_samp. The SSL standard method here is 4A1d if you want to look it up.
+ *bd_other* is, for data contributed by NRCS SSL, the bulk density of the fine earth fraction, but expressed on a field moist condition. It is collected by a volumetric core, but given the presence of some water, will probably usually be an overestimate relative to the bd_samp. This corresponds to the SSL standard method 4A3a. For datasets contributed by sources other than NRCS, bd_other is anyone's guess. Hopefully specified in the associated bd_method variable, or the metadata sheet for that dataset.

Reasonable BD minimum 0.03 or lower and maximum 2.7 g cm-3. 
```{r BD}
plot.df <- merge(subset(ISCN3$sample, grepl('^bd_', measurementID) & 
                          value > 0 & value < 2.7), ISCN3$measurement)
ggplot(plot.df) +
  geom_histogram(aes(x=value, y=..density..)) + 
  #scale_x_log10() + 
  facet_wrap(type~unit)

summary(subset(ISCN3$sample, grepl('^bd_', measurementID) & 
                          (value <= 0 | value > 2.7)))

nrow(plot.df)
```

Number of samples for BD `r length(plot.df$value)`.

## Carbon (mostly organic)

+ *c_tot* carbon concentration in a dry combustion analysis, sometimes include inorganic carbon
+ *oc* organic carbon concentration from either acidified (to remove carbonates) and then ran on dry combustion analyizer, or (more common) wet oxidation 'Walkly-Black' dicromate oxidation, chemical digest for organic carbon; old methods 1930-1990s tends to over estimate it in certain soils. pH below 7 c_tot generally == oc otherwise you might have carbonates
+ *loi* loss on ignition uncorrected for C:organics, generally close to half for O-horizon, C about a third of the LOI for lower horizons.

```{r C_OC_LOI}
ggplot(merge(subset(ISCN3$sample, grepl('^(c_|oc|loi|soc)', measurementID) & value > 0),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit)

summary(subset(ISCN3$sample, grepl('^(c_|oc|loi)', measurementID) & (value < 0)))
```

## Carbon (inorganic)
Percent inorganic carbon in a sample.

```{r CaCO3}
ggplot(merge(subset(ISCN3$sample, grepl('^caco3', measurementID) & value > 0),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit)

summary(subset(ISCN3$sample, grepl('^caco3', measurementID) & (value <= 0)))
```

## Nitrogen
+ *n_tot* most N is organic mass percent dry weight, probably estimated by dry combustion (Dumas methods), sometimes also Kjeldahl digestion converts N to nitrate and then measures nitrate.

```{r N}
ggplot(merge(subset(ISCN3$sample, grepl('^n_', measurementID) & value > 0),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit)
```

## pH

+ *ph_cacl* 1:2 Soil-CaCl2 suspension
+ *ph_h2o* 1:1 Soil-water suspension
+ *ph_other* non-standard catch all

```{r pH}
ggplot(merge(subset(ISCN3$sample, grepl('^ph_', measurementID) & 
                      value > 1.5 & value < 11),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  #scale_x_log10() + 
  facet_wrap(type~unit)

summary(subset(ISCN3$sample, grepl('^ph_', measurementID) & (value <= 1.5 | value > 11)))
#d_ply(subset(ISCN3$sample, grepl('^ph_', measurementID) & (value <= 1.5 | value > 11)), 
#      'measurementID', function(xx){print(summary(xx))})
```

## Sand, clay, slit, and percent coarse fragement content

Partical size analysis. 
Methods might not be completely annotated.

+ *wpg2* coarse fragment (>2mm) content by weight
+ *sand_tot_psa* percentage of sand
+ *clay_tot_psa* percentage of clay
+ *silt_tot_psa* percentage of silt

```{r Texture}
ggplot(merge(subset(ISCN3$sample, grepl('(wpg|tot_psa)', measurementID) & 
                      value < 100 & value >= 0 ),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  #scale_x_log10() + 
  facet_wrap(type~unit)

summary(subset(ISCN3$sample, grepl('(wpg|tot_psa)', measurementID) & 
                 (value < 0 | value > 100)))
```

##Cation exchange capasity and base saturation

+ *cat_exch* cation exchange capasity [[TODO: Ask Luke N about this variable]]
+ *cec_sum* cation exchange capsity (sum of cations)
+ *ecec* effective cation exchange capsity
+ *bs* base saturation (NH4OAc at pH 7.0)
+ *bs_sum* base saturation (sum of cations at pH 8.2)

```{r cat}
ggplot(merge(subset(ISCN3$sample, grepl('(cat_exch|cec_sum|ecec)', measurementID)& (value > 0)),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit)

summary(subset(ISCN3$sample, grepl('(cat_exch|cec_sum|ecec)', measurementID) & (value <= 0)))
```

```{r bs}
ggplot(merge(subset(ISCN3$sample, grepl('(bs|bs_sum)', measurementID)& (value > 0)),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit)

summary(subset(ISCN3$sample, grepl('(bs|bs_sum)', measurementID) & (value <= 0)))
```

## Extractable bases
+ *ca_al* TODO Luke N, what is this??
+ *ca_ext* calcium extractable (SSL: NH4OAc extractable)
+ *k_ext* potassium extractable (SSL: NH4OAc extractable)
+ *mg_ext* magnesium extractable (SSL: NH4OAc extractable)
+ *na_ext* sodium extractable (SSL: NH4OAc extractable)
+ *base_sum* sum of NH4OAc extractable bases
```{r metals_ext}
ggplot(merge(subset(ISCN3$sample, grepl('(ca|k|mg|na)', measurementID) & value > 0),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit, scales='free')

summary(subset(ISCN3$sample, grepl('(ca|k|mg|na)', measurementID) & (value <= 0)))
```

## Metal (Al, Fe, Mn) extracts
For ISCN3 this is only for NRCS data.

+ *_dith* Dithionite citrate extractable
+ *_ox* Oxalate extractable
+ *al_other* and *fe_other* Sodium phyrophospate extractable 
+ *mn_other* KCl extractable

```{r metals_percent}
ggplot(merge(subset(ISCN3$sample, 
                    grepl('^(al|fe|mn)_', measurementID) &
                      (value <= 100 & value >= 0)),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit, scales='free')

summary(subset(ISCN3$sample, grepl('^(al|fe|mn)_', measurementID) &
                      !grepl('ext', measurementID) & (value < 0 | value > 100)))
```

## Phosphorus
+ *p_bray* bray-1 extractable
+ *p_ox* oxalate extractable
+ *p_other* mehlich3 extractable

```{r p_percent}
ggplot(merge(subset(ISCN3$sample, grepl('^p_', measurementID)),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  scale_x_log10() + 
  facet_wrap(type~unit, scales='free')

summary(subset(ISCN3$sample, grepl('^p_', measurementID) & (value > 100 | value < 0)))
```

## Isotope

```{r isotope}
#15n          13c          14c
ggplot(merge(subset(ISCN3$sample, grepl('^1[345]', measurementID)),
             ISCN3$measurement)) +
  geom_histogram(aes(x=value, y=..density..)) + 
  #scale_x_log10() + 
  facet_wrap(~type, scales='free')

summary(subset(ISCN3$sample, grepl('^1[345]', measurementID)))
```