---
title: "Data ingestion for 2017"
author: "K Todd-Brown (ktoddbrown@gmail.com)"
date: "8 March 2018"
output:
  html_document: default
  pdf_document:
    fig_caption: yes
    toc: yes
---

```{r setup}
library(SoilDataR) #library(devtools); install_github("ISCN/soilDataR")
library(tidyverse)

#mapping librarys to help with global/regional plots
library(ggmap)
library(maps)
library(mapdata)
library(fiftystater)

```

Datasets are:
1) Treat: Peat properties synthesis dataset (2MB, XLSX format, download only; ISCNtemplate_Treat_peatProps_v2): This dataset is a synthesis of literature and site-level data on peat properties, C, N, 14C, and vegetation from 366 sites worldwide. Data are available for nearly 16,000 layers from 659 profiles. Data contributed by Claire Treat.
2) Alamos soil C stocks (<1MB, XLSX format, download only; ISCNtemplate_Alamos): This site-level dataset comes courtesy of Kris Johnson and collaborators at ITSON (Obregon, MX). It contains 30 profiles sampled by quantitative pit as part of a NASA-supported C monitoring study.

```{r download}
dataDir <- tempdir() #if you want to download to a temp directory
#dataDir <- '~/Documents/Datasets/temp' #rename if you prefer a different local director

TreatFilename <- file.path(dataDir, 'ISCNtemplate_Treat_peatProps_v2.xlsx')
if(!file.exists(TreatFilename)){
  download.file(
    'http://iscn.fluxdata.org/wp-content/uploads/sites/15/ISCNtemplate_Treat_peatProps_v2.xlsx', 
    TreatFilename, quiet=FALSE)
}

AlamosFilename <- file.path(dataDir,'ISCNtemplate_Alamos.xlsx')
if(!file.exists(AlamosFilename)){
  download.file(
    'http://iscn.fluxdata.org/wp-content/uploads/sites/15/ISCNtemplate_Alamos.xlsx', 
    AlamosFilename, quiet=FALSE)
}
##Alamose has some formatting issues so let's deal with that first
AlamosDownload <- AlamosFilename
AlamosFilename <- file.path(dataDir, c('ISCNtemplate_Alamos_metadata.csv',
                                       'ISCNtemplate_Alamos_site.csv',
                                       'ISCNtemplate_Alamos_profile.csv',
                                       'ISCNtemplate_Alamos_layer.csv'))
temp <-  readxl::read_excel(AlamosDownload, sheet='metadata')
readr::write_csv(path=AlamosFilename[1],
                 x=temp) #metadata sheet is fine
readr::write_csv(path=AlamosFilename[2],
                 x= readxl::read_excel(AlamosDownload, sheet='site')[-3,]) #remove empty row
readr::write_csv(path=AlamosFilename[3],
                 readxl::read_excel(AlamosDownload, sheet='profile')[-3,]) #remove empty row
readr::write_csv(path=AlamosFilename[4],
                 #remove empty row
                 readxl::read_excel(AlamosDownload, sheet='layer')[-3,] %>% 
                   #and fill in rows with non-NA values above
                   tidyr::fill(site_name) %>%
                   #make the layer name unique
                   mutate(layer_name = paste(profile_name, layer_name, sep='-')))

```

```{r readKey}
key.df <- read_csv('../dataset_keys/ISCNtemplate_2016Key.csv')
```

```{r loadData}

Treat <- SoilDataR::readKeyedData(filename=TreatFilename, key.df=key.df, dropRows=1:2)
Treat$long <- Treat$key %>% 
  filter(!is.na(hardUnit)) %>% #units are the only thing hard coded here
  select(var, hardUnit) %>% unique %>% 
  right_join(Treat$long, by=c('var')) %>% 
  rename(unit=hardUnit)

Alamos <- SoilDataR::readKeyedData(filename=AlamosFilename, 
                                   verticalTable = c(AlamosFilename[1]),
                                   key.df=key.df, dropRows=1:2)

Alamos$long <- Alamos$key %>% 
  filter(!is.na(hardUnit)) %>% #units are the only thing hard coded here
  select(var, hardUnit) %>% unique %>% 
  right_join(Alamos$long, by=c('var')) %>% 
  rename(unit=hardUnit)
```

```{r constructTables}
data.ls <- list()
data.ls$study <- Alamos$wide %>% bind_rows(Treat$wide) %>% #merge two datasets
  select_(.dots=intersect(names(.), c((Alamos$key %>% filter(dataframe == 'study'))$var,
           (Treat$key %>% filter(dataframe == 'study'))$var))) %>% #select vars
  unique %>%
  group_by_(.dots=unique(c((Alamos$key %>% 
                              filter(dataframe == 'study', !is.na(flagID)))$var,
           (Treat$key %>% 
              filter(dataframe == 'study', !is.na(flagID)))$var))) #group by key

data.ls$field <- Alamos$wide %>% bind_rows(Treat$wide) %>% #merge two datasets
  select_(.dots=intersect(names(.), 
                          c((Alamos$key %>% filter(dataframe == 'field' |
                                                     !is.na(flagID)))$var,
                            (Treat$key %>% filter(dataframe == 'field'|
                                                    !is.na(flagID)))$var))) %>% #select vars
  unique %>%
  mutate_at(vars('layer_top', 'layer_bottom', 'lon', 'lat'), funs(as.numeric)) %>%
  mutate(observation_date = 
              if_else(grepl('/', observation_date),
                      lubridate::decimal_date( lubridate::mdy(observation_date)),
                      as.numeric(observation_date))) %>%
  mutate_at(vars(-lon, -lat, -layer_top, -layer_bottom, -observation_date),
            funs(factor)) %>%
              group_by(dataset_name, layer_name, profile_name, site_name) #group by names
  
data.ls$sample <- Alamos$long %>% 
  mutate(dataset_name=unique(Alamos$wide$dataset_name)) %>%
  bind_rows(Treat$long %>% 
               mutate(dataset_name=unique(Treat$wide$dataset_name))) %>%
  ungroup() %>%
  mutate_at(vars(-value, -sigma),funs(factor)) %>%
  mutate_at(vars('value', 'sigma'), funs(as.numeric)) %>%
  group_by_(.dots=groups(data.ls$field))
```

#Observation time
```{r obsTime}
location.df <- data.ls$field %>% ungroup %>% 
  select(lat, lon, observation_date, state, country, dataset_name) %>%
  ###segment everything
  mutate(yrCut = cut(observation_date,
                     seq(from = 1960, 
                         to = 2020, by=20),
                     dig.lab=4),
         latCut = cut(lat, seq(-90, 90, by=0.05)),
         lonCut = cut(lon, seq(-180, 180, by=0.05)))

timeSpaceCounts <- location.df %>%
  group_by(yrCut,  country) %>%
  tally %>%
  spread(key=yrCut, value=n)

print(timeSpaceCounts)
```
```{r plotTime}
ggplot(location.df, aes(x=observation_date, fill=dataset_name)) + geom_histogram()

mapWorld <- borders("world", colour="gray80", fill="gray80") # create a layer of borders
#ggplot() + mapWorld
ggplot(location.df) +
  mapWorld + 
  #geom_hex(aes(x=long, y=lat), bins=200) + 
  geom_point(aes(x=lon, y=lat, color=dataset_name)) +
  scale_fill_gradient(trans='log10') +
  theme_bw() +
  theme(text=element_text(size=18),
        legend.text=element_text(size=10),
        axis.title=element_blank()) +
  #ylim(45, 90) + 
  #coord_map(projection='azequidistant') +
  facet_wrap(~yrCut)

```

# Site locations
## Lat-lon map
```{r mapSites}
mapWorld <- borders("world", colour="gray80", fill="gray80") # create a layer of borders
#ggplot() + mapWorld
ggplot(unique(location.df[, c('lat', 'lon')])) +
  mapWorld + 
  geom_hex(aes(x=lon, y=lat), bins=200) + 
  scale_fill_gradient(trans='log10') +
  theme_bw() +
  theme(text=element_text(size=18)) +
  labs(x='', y='')
```

# Measruements distribution
```{r measurementHist, fig.height=10, fig.width=7}
ggplot(data.ls$sample) +
  geom_histogram(aes(x=value, fill=dataset_name)) +
  facet_wrap(var~unit, scales='free', ncol=2)
```
