---
title: "Scratch"
author: "K Todd-Brown (ktoddbrown@gmail.com)"
date: "3/8/2018"
output: html_document
---

```{r setup}
library(SoilDataR) #library(devtools); install_github("ISCN/soilDataR")
library(tidyverse)

#mapping librarys to help with global/regional plots
library(ggmap)
library(maps)
library(mapdata)
library(fiftystater)

#dataDir <- tempdir() #if you want to download to a temp directory
dataDir <- '~/Documents/Datasets/temp' #rename if you prefer a different local director
```

```{r Moore2015}
downloadURL <- 'https://datadryad.org/bitstream/handle/10255/dryad.97546/Moore%202015%20Final%20Data.csv?sequence=1'


MooreFilename <- file.path(dataDir, 'Moore 2015 Final Data.csv')
if(!file.exists(MooreFilename)){
  download.file(downloadURL,  MooreFilename, quiet=FALSE)
}

##the permille ascii doesn't play nice so remove it from the header and key
temp <- read_csv(MooreFilename, col_names =TRUE)
names(temp)[7] <- "13C in resp (permille)"
write_csv(temp, path=MooreFilename)

key.df <- read_csv('../dataset_keys/Moore2015.csv')
key.df$header[grepl('13C in resp', key.df$header)] <- '13C in resp (permille)'

###read the file
Moore <- readKeyedData(filename=MooreFilename, key.df=key.df)

Moore$long <- Moore$key %>% 
  filter(!is.na(softType)) %>%
  filter(!is.na(hardUnit) | !is.na(hardMethod)) %>%
  rename(unit=hardUnit, method=hardMethod) %>%
  select(var, unit, method) %>%
  full_join(Moore$long)

Moore$wide <- Moore$key %>% 
  filter(is.na(softType)) %>%
  filter(!is.na(hardUnit) | !is.na(hardMethod) | !is.na(hardValue) | !is.na(hardSigma)) %>%
  select(var, contains('hard')) %>%
  gather(key='type', value='text', contains('hard'), na.rm=TRUE) %>%
  mutate(type = gsub('hard', '', type)) %>%
  mutate(header = paste(var, if_else(type == 'Value', '', paste('_', type, sep='')), sep='')) %>%
  select(header, text) %>%
  spread(key=header, value=text) %>%
  data.frame(Moore$wide)
```

#Shaw et al (in press)
Extremely well documented.

Shaw, C., Hilger, A., Filiatrault, M. and Kurz, W. (in press), A Canadian upland forest soil profile and carbon stocks database. Ecology. Accepted Author Manuscript. doi:10.1002/ecy.2159

Copyright restrictions
© Her Majesty the Queen in Right of Canada, 2017

Information contained in this publication or product may be reproduced, in part or in whole, and by any means, for personal or public non-commercial purposes, without charge or further permission, unless otherwise specified.
You are asked to:
-  exercise due diligence in ensuring the accuracy of the materials reproduced;
-  indicate the complete title of the materials reproduced, and the name of the author organization; and
-  indicate that the reproduction is a copy of an official work that is published by Natural Resources Canada (NRCan) and that the reproduction has not been produced in affiliation with, or with the endorsement of, NRCan.
-  Commercial reproduction and distribution is prohibited except with written permission from NRCan. For more information, contact NRCan at copyright.droitdauteur@nrcan-rncan.gc.ca.

```{r Shaw2018}
ShawDownload_url <- c('http://binarystore.wiley.com/store/10.1002/ecy.2159/asset/supinfo/ecy2159-sup-0002-MetadataS1.pdf?v=1&s=1b267bc2e7bbafc52b12a9e522033fa590445713', 
   'http://binarystore.wiley.com/store/10.1002/ecy.2159/asset/supinfo/ecy2159-sup-0001-DataS1.zip?v=1&s=f95f40084248b880723282404305327c0aa95aeb')


ShawDownloadFilename <- file.path(dataDir, 
                                  c('ecy2159-sup-0002-MetadataS1.pdf',
                                    'ecy2159-sup-0001-DataS1.zip'))
if(!all(file.exists(ShawDownloadFilename))){
  download.file(ShawDownload_url[1], ShawDownloadFilename[1], quiet=FALSE)
  download.file(ShawDownload_url[2], ShawDownloadFilename[2], quiet=FALSE)
}
ShawDataFiles <- file.path(file.path(dataDir, 'Shaw2018'),
                           unzip(ShawDownloadFilename[2], list=TRUE))

unzip(ShawDownloadFilename[2], exdir=file.path(dataDir, 'Shaw2018'))

key.df <- read_csv('../dataset_keys/Shaw2018.csv')
Shaw <- SoilDataR::readKeyedData(
  filename=file.path(dataDir, 'Shaw2018',
                     c('REFERENCES.csv', 'SITES.csv', 'PROFILES.csv')), key.df=key.df)

Shaw$long <- Shaw$key %>% ungroup() %>%
  filter(!is.na(softType)) %>%
  filter(!is.na(hardUnit) | !is.na(hardMethod)) %>%
  select(var, hardUnit, hardMethod) %>% ##only units and methods are hard in Shaw2018
  full_join(Shaw$long, by='var') %>% ungroup() %>%
  mutate(method=ifelse(is.na(hardMethod), method, hardMethod),
         unit=hardUnit) %>% #no soft units
  mutate(var=factor(var), 
         method=factor(method), 
         unit=factor(unit),
         value=as.numeric(value)) %>%
  select(-hardUnit, -hardMethod)

ggplot(Shaw$wide %>% select(lat, lon, observation_date) %>%
  mutate_all(funs(as.numeric))) +
  geom_histogram(aes(observation_date))

mapWorld <- borders("world", colour="gray80", fill="gray80") # create a layer of borders
#ggplot() + mapWorld
ggplot(Shaw$wide %>% 
         select(lat, lon, observation_date) %>%
         mutate_all(funs(as.numeric)) %>%
         mutate(decade = floor(observation_date/10)*10)) +
  mapWorld + 
  geom_hex(aes(x=lon, y=lat), bins=50) + 
  #geom_point(aes(x=lon, y=lat)) +
  theme_bw() +
  theme(text=element_text(size=18),
        legend.text=element_text(size=10),
        axis.title=element_blank()) +
  facet_wrap(~decade)

ggplot(Shaw$long) +
  geom_histogram(aes(x=value)) +
  facet_wrap(~var+unit, scale='free')
```
