---
title: "ISCN3 visualization and report"
author: "K Todd-Brown (ktoddbrown@gmail.com)"
date: "11/13/2017"
output:
  pdf_document:
    fig_caption: yes
    toc: yes
  html_document: default
---

The International Soil Carbon Network version 3 data base (ISCN3) is available [here](http://iscn.fluxdata.org/).
Currently the Layers data is ingestion and summarized below (work is underway to also bring in profile level data). 
Much of the data is from [NRCS](https://www.nrcs.usda.gov/wps/portal/nrcs/site/soils/home/) (Sept 2014 version of the soil characterization data base) and methods can be refered to in SSL (SoiL Survey Manual, R Burt) manual. 

```{r setup}
library(SoilDataR) #library(devtools); install_github("ktoddbrown/soilDataR")
library(ggplot2) #make pretty plots
library(lubridate)
library(dplyr)
library(tidyr)
library(readr)

#mapping librarys to help with global/regional plots
library(ggmap)
library(maps)
library(mapdata)
library(fiftystater)
```

```{r loadData}
ISCN3 <- processData_ISCN3(dataDir='~/Documents/Datasets/ISCN_3/', verbose=TRUE)
save(ISCN3, file='~/Documents/Datasets/ISCN_3/ISCN3.RData')
if(file.exists('../temp/ISCN3.RData')){
  load('../temp/ISCN3.RData')
}else{
  ISCN3 <- processData_ISCN3(verbose=TRUE)
  save(ISCN3, file='../temp/ISCN3.RData')
}
```
```{r cleanUp}
#load('../temp/ISCN3.RData')
load('~/Documents/Datasets/ISCN_3/ISCN3.RData')

##Add new indecies for possible duplications
ISCN3$field <- ISCN3$field %>% 
  mutate(fieldID_byString = group_indices(., lat, lon, 
                                        layer_bottom, layer_top, observation_date)) %>%
  mutate_at(c('layer_bottom', 'layer_top', 'observation_date'),
            funs(num=round(as.numeric(as.character(.)), 0))) %>%
  mutate_at(c('lat', 'lon'), funs(num=round(as.numeric(as.character(.)), 1))) %>%
  mutate(fieldID_byApprox = group_indices(., lat_num, lon_num, 
                                         layer_bottom_num, layer_top_num,
                                   observation_date_num)) %>%
  select(-ends_with('_num'))

##Identify non-located entries
noLocation <- ISCN3$field %>%
  filter(is.na(lat), is.na(lon), 
         is.na(layer_bottom), is.na(layer_top),
         is.na(observation_date)) %>%
  select(ends_with('_name'), starts_with('fieldID')) %>%
  unique()

##Identify conflicting entries
problematicFieldEntries <- ISCN3$field %>%
  filter(!is.na(lat), !is.na(lon), !is.na(layer_bottom), !is.na(layer_top),
         !is.na(observation_date)) %>% #only look at well defined sites
  select(-lat, -lon, -layer_bottom, -layer_top, -observation_date, 
         -fieldID, -fieldID_byApprox) %>% #look at non-identifying columns
  group_by(fieldID_byString) %>% #group by entry matching
  gather(key='header', value='value', -fieldID_byString, na.rm=TRUE) %>%
  group_by(fieldID_byString, header) %>%
  unique() %>%
  tally()  %>% #count up number of unique values for that header
  filter(n > 1) #only care about headers with more then one value
  
#ISCN3$field$fieldID_byString %>% unique %>% length
#problematicFieldEntries$fieldID_byString %>% unique %>% length
#problematicFieldEntries %>% arrange(-n) %>% head

##recast the lat-lon as numerics
ISCN3$field <- ISCN3$field  %>%
  mutate(lat=as.numeric(lat), lon=as.numeric(lon))

##Add the hard coded units
hardUnits.df <- ISCN3$ISCNKey %>% filter(type == 'value', !is.na(hardUnit)) %>% select(var, hardUnit)
ISCN3$measure <-  ISCN3$measure %>% 
  left_join(select(filter(ISCN3$ISCNKey, type=='value'), var, hardUnit), by='var') %>%
  mutate(unit=if_else(grepl('\\w', unit) , as.character(unit), hardUnit))

##Check country based on lat/lon

```


#Observation time
```{r obsTime}
location.df <- unique(ISCN3$field[, c('lat', 'lon', 'observation_date', 
                                      'state', 'country')]) %>%
  mutate(observation_date = as.numeric(as.character(observation_date)),
         lat = as.numeric(lat), lon=as.numeric(lon)) %>%
  mutate(observation_date=if_else(observation_date < 1900, as.numeric(NA), observation_date)) %>%
  #separate(observation_date, c('monthStr', 'dayStr', 'yearStr'), 
  #         remove=FALSE, fill='left') %>%
  #mutate(year=if_else(is.na(as.numeric(yearStr)), as.numeric(yearStr),
  #                    if_else(as.numeric(yearStr) < 20, as.numeric(yearStr) + 2000,
  #                        if_else(as.numeric(yearStr) < 100, as.numeric(yearStr) + 1900,
  #                                as.numeric(yearStr)))),
  #       month=if_else(is.na(as.numeric(monthStr)), 1,  as.numeric(monthStr)),
  #       day=if_else(is.na(as.numeric(dayStr)), 1, as.numeric(dayStr))) %>%
  #select(-ends_with('Str')) %>% 
  #mutate(obsDate = ymd(paste(year, month, day, sep='-'))) %>%
  mutate(year = floor(observation_date),
    obsDate = date_decimal(observation_date)) %>%
  arrange(lat, lon, obsDate) %>%
  mutate(yrCut = cut(year, seq(from = floor(min(year, na.rm=TRUE)/10)*10, 
                               to = ceiling(max(year, na.rm=TRUE)/10)*10, by=10),
                     dig.lab=4),
         latCut = cut(lat, seq(-90, 90, by=0.5)),
         lonCut = cut(lon, seq(-180, 180, by=0.5)))

timeSpaceCounts <- location.df %>%
  mutate(inUSA = country %in% 'United States') %>%
  group_by(yrCut, latCut, lonCut, inUSA) %>%
  tally
print(location.df %>%
  group_by(yrCut) %>% tally)
```
```{r plotTime}
ggplot(location.df, aes(x=obsDate)) + geom_histogram()

mapWorld <- borders("world", colour="gray80", fill="gray80") # create a layer of borders
#ggplot() + mapWorld
ggplot(location.df) +
  mapWorld + 
  geom_hex(aes(x=lon, y=lat), bins=50) + 
  scale_fill_gradient(trans='log10') +
  theme_bw() +
  theme(text=element_text(size=18), 
        axis.title.x = element_blank(), axis.title.y=element_blank()) +
  facet_wrap(~yrCut)

#ggplot(location.df, aes(x=long, y=lat)) + geom_point(alpha=0.1)

ggplot(timeSpaceCounts, aes(x=n)) + geom_histogram() + 
  scale_x_log10() + scale_y_log10()+ facet_wrap(~inUSA, scales='free') +
  labs(title='Close space-time sites by USA status')
```

# Site locations
## Lat-lon map
```{r mapSites}

mapWorld <- borders("world", colour="gray80", fill="gray80") # create a layer of borders
#ggplot() + mapWorld
sites <- location.df %>% select(lat, lon) %>% unique()
ggplot(sites) +
  mapWorld + 
  geom_hex(aes(x=lon, y=lat), bins=50) + 
  scale_fill_gradient(trans='log10') +
  theme_bw() +
  theme(text=element_text(size=18),
        axis.title.x = element_blank(), axis.title.y=element_blank())
```

## Sites with US states but no lat-lon
```{r statemap}
stateCounts <- ISCN3$field %>%
  filter(is.na(lat+lon)) %>%
  group_by(state, country) %>%
  tally %>%
  ungroup() %>%
  mutate(state=tolower(state),
         knownState = tolower(state) %in% fifty_states$id)

ggplot(stateCounts, aes(map_id = state)) +
  geom_map(aes(fill=n), map=fifty_states) +
  expand_limits(x = fifty_states$lon, y = fifty_states$lat) +
  coord_map() +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  scale_fill_gradient(trans='log10') +
  labs(x = "", y = "") +
  theme(legend.position = "bottom", 
        panel.background = element_blank())
```

## Sites with country but no lat-lon
```{r countryMap}
countryCounts <- stateCounts %>%
  group_by(country) %>%
  summarize(count=sum(n)) %>%
  mutate(region=if_else(grepl('United States', as.character(country)), 'USA',
                        as.character(country)))
  
map.world <- map_data(map="world")

cat('These countries are unknown to the map\n')
print(setdiff(countryCounts$region, map.world$region))

countryCounts <- countryCounts %>%
  bind_rows(data.frame(region = setdiff(map.world$region, countryCounts$region),
                       count = 0))

ggplot(countryCounts, aes(map_id = region)) +
  geom_map(aes(fill=count), map=map.world) +
  expand_limits(x = map.world$lon, y = map.world$lat) +
  coord_cartesian(xlim=c(-180,180), ylim=c(-90,90)) +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  scale_fill_gradient(trans='log10') +
  theme(legend.position = "bottom", 
        panel.background = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank())
```

`r sum(filter(countryCounts, (country=='Unknown' | country == ''))$count)` sites with no known country nor lat-lon.

## Site count by country map (include sites w/ lat-lon)
```{r totCountryCount}
countryCounts <- ISCN3$field %>%
  filter(!is.na(lat+lon)) %>%
  group_by(country) %>%
  tally %>%
  mutate(region=if_else(grepl('United States', as.character(country)), 'USA',
                        as.character(country)))
  

map.world <- map_data(map="world")

countryCounts <- countryCounts %>%
  bind_rows(data.frame(region = setdiff(map.world$region, countryCounts$region),
                       n = 0))

ggplot(countryCounts, aes(map_id = region)) +
  geom_map(aes(fill=n), map=map.world) +
  expand_limits(x = map.world$lon, y = map.world$lat) +
  coord_cartesian(xlim=c(-180,180), ylim=c(-90,90)) +
  scale_x_continuous(breaks = NULL) + 
  scale_y_continuous(breaks = NULL) +
  scale_fill_gradient(trans='log10') +
  theme(legend.position = "bottom", 
        panel.background = element_blank(),
        axis.title.x=element_blank(),
        axis.title.y=element_blank())
```

# Summary of measures

The following variables are listed in 
```{r types}
unique(ISCN3$measure$var)
```
All measurements are on oven dry soils and in mass percent where percent noted.

## Soil organic carbon stock

+ *soc* calculated soil carbon stock of layer/profile

```{r SOC}
pullVarVis.fun <- function(var.regex='^soc', value.filter=function(xx){xx > 0}){
  plot.df <- inner_join(ISCN3$sample, 
                        ISCN3$measure %>% filter(grepl(var.regex, var)))
  myPlot <- ggplot(plot.df %>% filter(value.filter(value))) +
    geom_histogram(aes(x=value, y=..density..)) +
    facet_wrap(var~unit)
  
  mySummary <- summary(plot.df %>% filter(!value.filter(value)))
  return(list(plot=myPlot, neg.summmary=mySummary))
}

printMe <- pullVarVis.fun(var.regex='^soc', value.filter=function(xx){xx > 0})
print(printMe$plot + scale_x_log10())
print(printMe$neg.summmary)
```

## Bulk density

+ *bd_samp* is the bulk density of the <2mm (fine earth) fraction, in which the mass is expressed on an oven-dry (105 deg) basis. This is the best form of Db for the most uses.
+ *bd_tot* is the whole soil bulk density, includes fine earth fraction and rocks.
+ *bd_whole* is the bulk density of the whole soil (coarse frags, if present, included), expressed on a 1/3 bar moisture content (a little drier than typical field moist, but not oven-dry). It is collected by the clod method. For a variety of reasons, including the presence of water, possible coarse frags, and the tendency of clods to cleave along pore spaces, this will overestimate Db relative to bd_samp. The SSL standard method here is 4A1d if you want to look it up.
+ *bd_other* is, for data contributed by NRCS SSL, the bulk density of the fine earth fraction, but expressed on a field moist condition. It is collected by a volumetric core, but given the presence of some water, will probably usually be an overestimate relative to the bd_samp. This corresponds to the SSL standard method 4A3a. For datasets contributed by sources other than NRCS, bd_other is anyone's guess. Hopefully specified in the associated bd_method variable, or the metadata sheet for that dataset.

Reasonable BD minimum 0.03 or lower and maximum 2.7 g cm-3. 
```{r BD}
printMe <- pullVarVis.fun(var.regex='^bd_', value.filter=function(xx){xx > 0 & xx < 2.7})
print(printMe$plot)
print(printMe$neg.summmary)
```

## Carbon (mostly organic)

+ *c_tot* carbon concentration in a dry combustion analysis, sometimes include inorganic carbon
+ *oc* organic carbon concentration from either acidified (to remove carbonates) and then ran on dry combustion analyizer, or (more common) wet oxidation 'Walkly-Black' dicromate oxidation, chemical digest for organic carbon; old methods 1930-1990s tends to over estimate it in certain soils. pH below 7 c_tot generally == oc otherwise you might have carbonates
+ *loi* loss on ignition uncorrected for C:organics, generally close to half for O-horizon, C about a third of the LOI for lower horizons.

```{r C_OC_LOI}
printMe <- pullVarVis.fun(var.regex='^(c_|oc|loi|soc)', value.filter=function(xx){xx > 0})
print(printMe$plot + scale_x_log10())
print(printMe$neg.summmary)
```

### Bulk density vs organic (mostly) carbon

```{r OCvsBD}
printMe <- inner_join(ISCN3$sample, 
                        ISCN3$measure %>% filter(grepl('^(c_|oc|loi|bd_)', var)), by='measureID') %>%
  select(fieldID, var, value) %>%
  mutate(axis = if_else(grepl('^bd_', .$var), 'x', 'y'),
         priority = recode(var, bd_sample=1, bd_other=2, bd_tot=4, bd_whole=3,
                           c_tot=1, oc=2, loi=3)) %>% 
  group_by(fieldID, axis) %>%
  summarize(value=value[min(priority)], var=var[min(priority)], priority=min(priority)) %>%
  filter(is.finite(value))

temp <- printMe %>% filter(fieldID > 445043)
compare.df <- inner_join(printMe %>% select(fieldID, axis, value) %>% spread(axis, value), 
                         printMe %>% select(fieldID, axis, var) %>% spread(axis, var), by='fieldID', suffix=c('.value', '.var')) %>%
  filter(is.finite(x.value+y.value))

ggplot(compare.df) + geom_bin2d(aes(x=x.value, y=y.value)) + facet_grid(x.var~y.var, scales='free')

ggplot(compare.df) + geom_histogram(aes(x=x.value/100*y.value)) + facet_wrap(~x.var+y.var, scales='free') + scale_x_log10()
```

## Carbon (inorganic)
Percent inorganic carbon in a sample.

```{r CaCO3}
printMe <- pullVarVis.fun(var.regex='^caco3', value.filter=function(xx){xx > 0})
print(printMe$plot + scale_x_log10())
print(printMe$neg.summmary)
```

## Nitrogen
+ *n_tot* most N is organic mass percent dry weight, probably estimated by dry combustion (Dumas methods), sometimes also Kjeldahl digestion converts N to nitrate and then measures nitrate.

```{r N}
printMe <- pullVarVis.fun(var.regex='^n_', value.filter=function(xx){xx > 0})
print(printMe$plot + scale_x_log10())
print(printMe$neg.summmary)
```

## pH

+ *ph_cacl* 1:2 Soil-CaCl2 suspension
+ *ph_h2o* 1:1 Soil-water suspension
+ *ph_other* non-standard catch all

```{r pH}
printMe <- pullVarVis.fun(var.regex='^ph_', value.filter=function(xx){xx  > 1.5 & xx < 11})
print(printMe$plot)
print(printMe$neg.summmary)
```

## Sand, clay, slit, and percent coarse fragement content

Partical size analysis. 
Methods might not be completely annotated.

+ *wpg2* coarse fragment (>2mm) content by weight
+ *sand_tot_psa* percentage of sand
+ *clay_tot_psa* percentage of clay
+ *silt_tot_psa* percentage of silt

```{r Texture}
printMe <- pullVarVis.fun(var.regex='(wpg|tot_psa)', value.filter=function(xx){xx  >=0 & xx < 100})
print(printMe$plot)
print(printMe$neg.summmary)
```

##Cation exchange capasity and base saturation

+ *cat_exch* cation exchange capasity [[TODO: Ask Luke N about this variable]]
+ *cec_sum* cation exchange capsity (sum of cations)
+ *ecec* effective cation exchange capsity
+ *bs* base saturation (NH4OAc at pH 7.0)
+ *bs_sum* base saturation (sum of cations at pH 8.2)

```{r cat}
printMe <- pullVarVis.fun(var.regex='(cat_exch|cec_sum|ecec)', value.filter=function(xx){xx > 0})
print(printMe$plot + scale_x_log10())
print(printMe$neg.summmary)

```

```{r bs}
printMe <- pullVarVis.fun(var.regex='(bs|bs_sum)', value.filter=function(xx){xx > 0})
print(printMe$plot + scale_x_log10())
print(printMe$neg.summmary)
```

## Extractable bases
+ *ca_al* TODO Luke N, what is this??
+ *ca_ext* calcium extractable (SSL: NH4OAc extractable)
+ *k_ext* potassium extractable (SSL: NH4OAc extractable)
+ *mg_ext* magnesium extractable (SSL: NH4OAc extractable)
+ *na_ext* sodium extractable (SSL: NH4OAc extractable)
+ *base_sum* sum of NH4OAc extractable bases
```{r metals_ext}
printMe <- pullVarVis.fun(var.regex='^(ca|k|mg|na)', value.filter=function(xx){xx > 0})
print(printMe$plot + scale_x_log10())
print(printMe$neg.summmary)
```

## Metal (Al, Fe, Mn) extracts
For ISCN3 this is only for NRCS data.

+ *_dith* Dithionite citrate extractable
+ *_ox* Oxalate extractable
+ *al_other* and *fe_other* Sodium phyrophospate extractable 
+ *mn_other* KCl extractable

```{r metals_percent}
printMe <- pullVarVis.fun(var.regex='^(al|fe|mn)_', value.filter=function(xx){xx >= 0 & xx <= 100})
print(printMe$plot + scale_x_log10())
print(printMe$neg.summmary)
```

## Phosphorus
+ *p_bray* bray-1 extractable
+ *p_ox* oxalate extractable
+ *p_other* mehlich3 extractable

```{r p_percent}
printMe <- pullVarVis.fun(var.regex='^p_', value.filter=function(xx){xx >= 0 & xx <= 100})
print(printMe$plot + scale_x_log10())
print(printMe$neg.summmary)
```

## Isotope

```{r isotope}

printMe <- pullVarVis.fun(var.regex='1[345]', value.filter=function(xx){TRUE})
print(printMe$plot)
print(printMe$neg.summmary)
```